---
title: "Survey Package and Regression"
author: ""
date: "February 25, 2025 "
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Let's work through the following R code:

```{r}
library(tidyverse)
```

# Loading in Data

The `dpylr` package has some data about Star Wars characters. Let's
assume it is a representative sample of all characters seen in Episodes
1 to 9.

```{r, message=FALSE}

starwars<-starwars

glimpse(starwars)

head(starwars)

```

We will be working with `mass`, and `height` throughout this class, so
let's remove the NAs. Additionally, we will be working with `species`,
let's categorize species to be `human`, \`droid\` or `other`.

```{r}
summary(starwars$mass)
summary(starwars$height)
table(starwars$species)

starwars_clean <- starwars %>% 
  filter(!is.na(mass)) %>% 
  filter(!is.na(height)) %>%
  mutate(species_clean = case_when(
    species == "Human" ~ "human",
    species == "Droid" ~ "droid",
    species != "Human" & species != "Droid" ~ "other")) %>% 
  filter(!is.na(species_clean))

starwars_clean 

n=nrow(starwars_clean)
```

### Task 1: Create a Simple Linear Model to predict mass

Use the `svyglm()` function in the `survey` library to run a linear
regression estimation of mass given height. Assume it was a Simple
Random Sample and use the finite population correction with $N=224$.

```{r}
library(survey)

N=224
n=nrow(starwars_clean)

fpc.srs = rep(N, n)

starwars.design <- svydesign(id=~1, data=starwars_clean, fpc=fpc.srs)

mysvylm <- svyglm(mass ~ height, starwars.design)
summary(mysvylm)
```

What is the model in this example?

$$y_{mass} = \beta_0 + \beta_1 x_{height} + \epsilon$$

This is a "simple linear regression" model (because there is only one
numeric $x$ variable).

What is the estimate of the model?

$$\hat{y}_{mass} = \hat{\beta_0} + \hat{\beta_1} x_{height}$$
$$\hat{y}_{mass} = -11.25 + 0.63 x_{height}$$

### Task 2: Create a Linear Model to predict mass

Use the `lm()` function to run a linear regression estimation of mass
given height.

```{r}
## Using the Survey Library
summary(lm(mass ~ height, data=starwars_clean))

```

What is the model in this example?

$$y_{mass} = \beta_0 + \beta_1 x_{height} + \epsilon$$ This is a "simple
linear regression" model (because there is only one numeric $x$
variable).

What is the estimate of the model?

$$\hat{y}_{mass} = \hat{\beta_0} + \hat{\beta_1} x_{height}$$
$$\hat{y}_{mass} = -11.25 + 0.63 x_{height}$$

What is the different between the output here and the outcome in Task 1?
What is similar?

**Standard errors are different, but estimates are the same.**

What happens if you change the $N$ in Task 1? Try setting
$N=87, 224, 1000$

### Task 3: Create a Linear Model to predict mass

Run a linear regression estimation of mass given height and species.

```{r}
## Using the Survey Library
fpc.srs = rep(N, n)

starwars.design <- svydesign(id=~1, data=starwars_clean, fpc=fpc.srs)

mysvylm <- svyglm(mass ~ height + species_clean, starwars.design)
summary(mysvylm) ### SRS

## Using lm
mymodel<-lm(mass ~ height + species_clean, data=starwars_clean)
summary(mymodel) ## TAUGHT IN STA302

```

What is the model in this example?

$$y_{mass} = \beta_0 + \beta_1 x_{height} + \beta_2x_{human} + \beta_3x_{other} + \epsilon$$
This is a "multiple linear regression" model (because there is are two
(more than 1) predictor $x$ variables).

What is the estimate of the model?

$$\hat{y}_{mass} = \hat{\beta_0} + \hat{\beta_1} x_{height} + \hat{ \beta_2}x_{human} + \hat{\beta_3}x_{other}$$
$$\hat{y}_{mass} = -21.38 + 0.65 x_{height} -14.64x_{human} + 20.18x_{other}$$
Note, the variables $x_{human}$ and $x_{other}$ are "dummy" variables.
They are coded such that they indicate whether or not the character is
in that species category.

Based on the output, what is the estimated mass of a human character who
is 175 units tall?

```{r}
-21.3802493 + 0.6509304 *175 - 14.6399467*1 +20.1821107*0
predict(mysvylm, tibble(height=175, species_clean="human"))
predict(mymodel, tibble(height=175, species_clean="human"))
```

Based on the output, what is the estimated mass of a droid character who
is 175 units tall?

```{r}
-21.3802493 + 0.6509304 *175 - 14.6399467*0 +20.1821107*0
predict(mysvylm, tibble(height=175, species_clean="droid"))
predict(mymodel, tibble(height=175, species_clean="droid"))
```

What is the expected difference in the mass of a human vs a droid
character of the same height?

```         
77.89262-92.53256 = -14.63994
```

We expect human characters to have a mass of 14.64 units lower than
droid characters of the same height.

### Task 4: Create a Logistic Regression Model to predict if mass exceeds 100lb.

Run a logistic regression model mass being over 100lbs given height and
species.

```{r}

## Create a new variable called `mass_over100`. 

starwars_clean <- starwars_clean %>%  mutate(
  mass_over100 = case_when(
    mass > 100 ~ 1, ## 1 = Yes, over 100lbs
    mass <= 100 ~ 0))


## Using lm
my_glm <- glm(mass_over100 ~ height + species_clean, 
              data=starwars_clean, 
              family = "binomial")
summary(my_glm)

```

What is the model in this example? (A2 - # 3 Methods)

$$log(\frac{p}{1-p}) = \beta_0 + \beta_1 x_{height} + \beta_2x_{human} + \beta_3x_{other}$$

What is the estimate of the model? (A2 - # 4 Results)

$$log\Big(\frac{\hat{p}}{1-\hat{p}}\Big) = \hat{\beta_0} + \hat{\beta_1} x_{height} + \hat{ \beta_2}x_{human} + \hat{\beta_3}x_{other}$$
$$log\Big(\frac{\hat{p}}{1-\hat{p}}\Big) = -11.45 + 0.06 x_{height} -2.14x_{human} -2.04 x_{other}$$

Based off the above output, what is the predicted probability of a human
who is 175cm tall being over 100lbs?

```{r}
pred_odds <- -11.45547800 + 0.06230662*175 -2.14452294*1 -2.04236670*0 
exp(pred_odds)/(1+exp(pred_odds))

pred_odds2 <- predict(my_glm, tibble(species_clean="human", height=175))
exp(pred_odds2)/(1+exp(pred_odds2))

pred_prob <- predict(my_glm, tibble(species_clean="human", height=175), type = "response")
pred_prob
```

```{r}
starwars.design <- svydesign(id=~1, data=starwars_clean, fpc=fpc.srs)

mysvyglm <- svyglm(mass_over100 ~ height + species_clean, 
              family = "binomial", starwars.design)

summary(mysvyglm)
```

### Task 5: Create a logistic regression for Stratified Random Sampling (A2)

For the starwars data, let’s assume the study instead collected the data
by stratifying by species, where in the population there 16 droids, 80
humans and 128 other species. Use the finite population correction with
N=224 to estimate.

Run a logistic regression model mass being over 100lbs given height and
species.

```{r}
table(starwars_clean$species_clean) # Sample strata sizes (ni's) are 4, 20, 32, total sample size (n) = 4 + 20 + 32 = 56

starwars_clean <- starwars_clean %>% 
  mutate(pop_fpc = case_when(species_clean=="droid" ~ 16,         # Population strata sizes (Ni's) are 16, 80, 128
                             species_clean=="human" ~ 80,         # Total population size (N) = 16 + 80 + 128 = 224
                             species_clean=="other" ~ 128)) %>% 
  mutate(samp_wt = case_when(species_clean=="droid" ~ 16/4,       # Sampling weights (samp_wt) are calculated as Ni/ni
                             species_clean=="human" ~ 80/20,     
                             species_clean=="other" ~ 128/32))


strata.design <- svydesign(id=~1,
                       strata=~species_clean,
                       weights = ~samp_wt, 
                       fpc=~pop_fpc,
                       data=starwars_clean)

mysvyglm <- svyglm(mass_over100 ~ height + species_clean, 
              family = "binomial", strata.design)

summary(mysvyglm)
```

\newpage

# Logistic Regression UCLA analysis of SRS with population of N=6000 code

```{r}
mydata <- read.csv("https://stats.idre.ucla.edu/stat/data/binary.csv")

## Standard Logistic Regression
mylogit<-glm(admit ~ gre + gpa +
               as.factor(rank), data=mydata, family="binomial")
summary(mylogit)

## log(p/(1-p)) = -3.99 + 0.002*gre + 0.804*gpa-0.675*x3
##                           -1.34*x4 - 1.55*x5


## Survey Estimation for Logistic Regression
n=length(mydata$admit)
N=6000

#install.packages("survey")
library(survey)
## Using the Survey Library
fpc.srs = rep(N, n)

ucla.design <- svydesign(id=~1, data=mydata, fpc=fpc.srs)

mysvyglm <- svyglm(admit ~ gre + gpa + as.factor(rank), 
                   ucla.design, family="binomial")
summary(mysvyglm)

```
